### 2021-05-07 nih2/NIHGrant.tex
Draft to Adam

Reverse correlation, in addition to its prominence as an important tool relative to lower-level
neural mechanisms, has been increasingly used for inferring higher-level cognitive representations,
including psychophysical kernels that drive the top-down processes of perception
\cite{ahumadaStimulusFeaturesSignal1971, neriReceptivePerceptiveFields2006,gosselinSuperstitiousPerceptionsReveal2003,smithMeasuringInternalRepresentations2012}, 
and even abstract psychological
categories (\eg ``male'' vs. ``female'' faces; \cite{brinkmanVisualisingMentalRepresentations2017,manginiMakingIneffableExplicit2004,ponsotCrackingSocialCode2018}.
Recent work in vision has demonstrated that
reverse correlation can effectively characterize cognitive representations underlying letter and face
recognition \hl{(Fig 2A)} \cite{gosselinSuperstitiousPerceptionsReveal2003,liuSeeingJesusToast2014}.
One speech study has also estimated steady-state representations of vowels the /a/
and /i/ using a closely-related paradigm to the one proposed here (Fig 2B; \cite{brimijoinInternalRepresentationVowel2013}). Our
proposed use of reverse correlation in the domain of speech expands on classic efforts to understand
top-down processing in speech with the presentation of rich, vague stimuli to elicit responses from
listeners \hl{(Warren \& Warren, 1970; Vokey \& Read 1985)}. These studies provide clear evidence that rich
stimuli, such as white noise, are sufficient to engage the top-down process, even if they do not attempt to
characterize the latent representations.

@@

We propose to develop an advanced signal processing pipeline that will enable us to overcome
the inefficiencies of existing reverse correlation methods through the use of compressive sensing, a
recent advance in signal processing which has led to dramatic improvements the efficiency of traditional
sampling and signal estimation methods (Baraniuk, 2007). Compressive sensing has recently gained
wide recognition in domains such as medical imaging (Graff \& Sidky, 2015; Lustig et al., 2008), where
considerations of efficiency and bias reduction are critical. Compressive sensing holds promise to
similarly improve the efficiency of reverse correlation, without the drawback of biasing estimates. By
dramatically decreasing the number of trials needed for signal reconstruction, this technique will extend
the range of perceptual mechanisms that can be estimated. Moreover, compressive sensing can be
directly substituted for conventional, regression-based estimation, with no other required changes to
existing experimental protocols. Our ultimate objective is to develop and validate a compressive sensing
data processing pipeline - culminating in an open-source software tool – that will allow for efficient and
accurate reconstruction of latent representations using data obtained via the reverse correlation method.

@@

The proposed project is innovative for several reasons.
First, we will leverage interdisciplinary collaboration between engineering (Lammert)
and hearing science (Ben) to characterize tinnitus to inform clinical care.
Second, we will apply reverse correlation to the tinnitus domain in a novel application,
which promises to deliver unbiased, high-dimensional estimates of cognitive representations of tinnitus.
Third, we will develop a signal processing pipeline to leverage compressive sensing
to acquire accurate estimates of cognitive representations of tinnitus using significantly fewer samples.
This pipeline will save experimenters and subjects time and will enable researchers
to collect data from more subjects and from individual subjects over time.
We will release our open-source software tool online with bindings to popular scientific computing languages.

@@

Our preliminary testing suggests that truly random envelopes
often result in the perception of colored noise.
Minimal additional spectral-temporal constraints may need to be imposed on stimuli
such that participants will interpret the stimuli as intending to be tinnitus percepts.
This is similar to the approach taken in some visual domain experiments,
where bandlimited random stimuli, or stimuli with partial superimposed templates,
are presented \cite{smithMeasuringInternalRepresentations2012} (\hl{CITE}).

\hl{What are we doing to create the stimuli?}

% \begin{description}
% 	\item[C.1.2. Experimental Procedure] 
% \end{description}

@@

\hl{Where are we getting our subjects?}

\hl{How are we evaluating tinnitus in our subjects?}

\hl{How do we phrase what we want the subjects to do?}

% \begin{description}
% 	\item[C.1.3. Reconstructing Representations] 
% \end{description}

@@

To reconstruct cognitive representations from subject responses,
we will generate a classification spectrogram by taking the difference
between a spectrogram generated from the positive labels (the ``yes'' responses)
and a spectrogram generated from the negative labels (the ``no'' responses)
(\autoref{fig:experiment}).
Such reconstructed representations are the best least-squares fit to the behavioral data
\cite{gosselinSuperstitiousPerceptionsReveal2003}.
We theorize that the classification spectrogram is used as a template vector
against which stimuli were compared in a linear matching model of response generation.

@@

The impact of this inefficiency extends well beyond the present proposal.
This fundamental sampling problem affects similar efforts to estimate higher-level
cognitive representations and psychophysical kernels that drive top-down processes of perception
\cite{ahumadaStimulusFeaturesSignal1971,neriReceptivePerceptiveFields2006,gosselinSuperstitiousPerceptionsReveal2003,smithMeasuringInternalRepresentations2012}.
Since reverse correlation has broad applicability for characterizing any aspect
of neurological, cognitive, or psychological function that can be modeled as a transductive process
\cite{ringachReverseCorrelationNeurophysiology2004},
the inefficiencies of reverse correlation impact efforts
to characterize phenomena as \hl{[word meaning minute, fine-grained, detailed, specific]} latent neural representations (\eg neural receptive fields) \cite{ringachReverseCorrelationNeurophysiology2004}
and as high-level  as abstract psychological categories (\eg ``male'' vs. ``female'' faces)
\cite{brinkmanVisualisingMentalRepresentations2017,ponsotCrackingSocialCode2018,manginiMakingIneffableExplicit2004}.
To combat this efficiency limitation, studies often 
(\emph{a}) limit the richness of the stimuli (\eg by only allowing specific aspects of the stimulus to vary)
\cite{gosselinBubblesTechniqueReveal2001},
or (\emph{b}) impose some constraints on the inferred representations,
for instance by smoothing the raw estimates \cite{gosselinSuperstitiousPerceptionsReveal2003}.
The constraints imposed by these approaches ultimately defeat the full power of reverse correlation
\cite{murrayTroublesBubbles2004},
leading to estimates that are biased or incomplete.

@@

% \subsubsection*{C.2.1. Method}

% In this section, we describe compressive sensing,
% including applicability to data collected in the reverse correlation paradigm,
% and demonstrate its effectiveness for improving the efficiency and accuracy
% of estimates of cognitive representations.

@@

\begin{enumerate}
	\item What is the shape of the manifold of cognitive representations of tinnitus?
	\item Is it reducible to fewer dimensions using PCA/t-SNE/UMAP?
	\item Can clusters on the manifold/reduced representation be found?
	What are those clusters?
	\item Is this insight clinically useful?
\end{enumerate}

>>>>

Reverse correlation, in addition to its prominence as an important tool relative to lower-level
neural mechanisms, has been increasingly used for inferring higher-level cognitive representations,
including psychophysical kernels that drive the top-down processes of perception
\cite{ahumadaStimulusFeaturesSignal1971, neriReceptivePerceptiveFields2006,gosselinSuperstitiousPerceptionsReveal2003,smithMeasuringInternalRepresentations2012}, 
and even abstract psychological
categories (\eg ``male'' vs. ``female'' faces; \cite{brinkmanVisualisingMentalRepresentations2017,manginiMakingIneffableExplicit2004,ponsotCrackingSocialCode2018}.
Recent work in vision has demonstrated that
reverse correlation can effectively characterize cognitive representations underlying letter and face
recognition \hl{(Fig 2A)} \cite{gosselinSuperstitiousPerceptionsReveal2003,liuSeeingJesusToast2014}.
One speech study has also estimated steady-state representations of vowels the /a/
and /i/ using a closely-related paradigm to the one proposed here (Fig 2B; \cite{brimijoinInternalRepresentationVowel2013}). Our
proposed use of reverse correlation in the domain of speech expands on classic efforts to understand
top-down processing in speech with the presentation of rich, vague stimuli to elicit responses from
listeners \cite{warrenAuditoryIllusionsConfusions1970,vokeySubliminalMessagesDevil1985}.
These studies provide clear evidence that rich
stimuli, such as white noise, are sufficient to engage the top-down process, even if they do not attempt to
characterize the latent representations.

@@

We propose to develop an advanced signal processing pipeline that will enable us to overcome
the inefficiencies of existing reverse correlation methods through the use of compressive sensing, a
recent advance in signal processing which has led to dramatic improvements the efficiency of traditional
sampling and signal estimation methods \cite{baraniukCompressiveSensingLecture2007}. Compressive sensing has recently gained
wide recognition in domains such as medical imaging \cite{graffCompressiveSensingMedical2015,lustigCompressedSensingMRI2008}, where
considerations of efficiency and bias reduction are critical. Compressive sensing holds promise to
similarly improve the efficiency of reverse correlation, without the drawback of biasing estimates. By
dramatically decreasing the number of trials needed for signal reconstruction, this technique will extend
the range of perceptual mechanisms that can be estimated. Moreover, compressive sensing can be
directly substituted for conventional, regression-based estimation, with no other required changes to
existing experimental protocols. Our ultimate objective is to develop and validate a compressive sensing
data processing pipeline - culminating in an open-source software tool – that will allow for efficient and
accurate reconstruction of latent representations using data obtained via the reverse correlation method.

@@

The proposed project is innovative for several reasons.
First, we will leverage interdisciplinary collaboration between engineering (Adam Lammert)
and hearing science (Ben Parrell) to characterize tinnitus to inform clinical care.
Second, we will apply reverse correlation to the tinnitus domain in a novel application,
which promises to deliver unbiased, high-dimensional estimates of cognitive representations of tinnitus.
Third, we will develop a signal processing pipeline to leverage compressive sensing
to acquire accurate estimates of cognitive representations of tinnitus using significantly fewer samples.
This pipeline will save experimenters and subjects time and will enable researchers
to collect data from more subjects and from individual subjects over time.
We will release our open-source software tool online with bindings to popular scientific computing languages.

@@

Our preliminary testing suggests that truly random envelopes
often result in the perception of colored noise.
Minimal additional spectral-temporal constraints may need to be imposed on stimuli
such that participants will interpret the stimuli as intending to be tinnitus percepts.
This is similar to the approach taken in some visual domain experiments,
where bandlimited random stimuli, or stimuli with partial superimposed templates,
are presented \cite{smithMeasuringInternalRepresentations2012} (\hl{CITE}).
The basic approach will be to use inverse Fourier transforms to generate 0.5-s noise stimuli
including frequencies from 0.1 to 22 kHz.
We will impose minimal constraints on the stimuli generation process,
in order to selectively generate stimuli approaching the heterogeneous samples of tinnitus percepts
provided by the American Tinnitus Association on their website.

@@

Subjects will be recruited from the clinical population by \hl{our clinical collaborator}
and will be assessed by the Tinnitus Severity Scale.
We will not screen for unilateral or bilateral tinnitus but will only include subjects
with primary (\ie subjective, not somatosensory) tinnitus.

During the experiment, subjects will be asked to perform an computerized alternate forced-choice task,
selecting whether the presented stimulus sounds similar to their tinnitus percept or not.
Each signal will be presented to subjects over headphones
at a comfortable listening level ($\approx 75$ dB SPL).
Subjects will be asked to perform the task in hour-long blocks with a rest period in-between.
Subjects can perform the task for one or more blocks at their choice,
and will be compensated for their participation.

@@

To reconstruct cognitive representations from subject responses,
we will generate a classification spectrogram by taking the difference
between a spectrogram generated from the positive labels (the ``yes'' responses)
and a spectrogram generated from the negative labels (the ``no'' responses)
(\autoref{fig:experiment}).
Such reconstructed representations are the best least-squares fit to the behavioral data
\cite{gosselinSuperstitiousPerceptionsReveal2003}.
We theorize that the classification spectrogram is used as a template vector
against which stimuli are compared in a linear matching model of response generation.

@@

The impact of this inefficiency extends well beyond the present proposal.
This fundamental sampling problem affects similar efforts to estimate higher-level
cognitive representations and psychophysical kernels that drive top-down processes of perception
\cite{ahumadaStimulusFeaturesSignal1971,neriReceptivePerceptiveFields2006,gosselinSuperstitiousPerceptionsReveal2003,smithMeasuringInternalRepresentations2012}.
Since reverse correlation has broad applicability for characterizing any aspect
of neurological, cognitive, or psychological function that can be modeled as a transductive process
\cite{ringachReverseCorrelationNeurophysiology2004},
the inefficiencies of reverse correlation impact efforts
to characterize phenomena as fine-grained as latent neural representations (\eg neural receptive fields) \cite{ringachReverseCorrelationNeurophysiology2004}
and as high-level  as abstract psychological categories (\eg ``male'' vs. ``female'' faces)
\cite{brinkmanVisualisingMentalRepresentations2017,ponsotCrackingSocialCode2018,manginiMakingIneffableExplicit2004}.
To combat this efficiency limitation, studies often 
(\emph{a}) limit the richness of the stimuli (\eg by only allowing specific aspects of the stimulus to vary)
\cite{gosselinBubblesTechniqueReveal2001},
or (\emph{b}) impose some constraints on the inferred representations,
for instance by smoothing the raw estimates \cite{gosselinSuperstitiousPerceptionsReveal2003}.
The constraints imposed by these approaches ultimately defeat the full power of reverse correlation
\cite{murrayTroublesBubbles2004},
leading to estimates that are biased or incomplete.

@@

@@

Our approach in this goal is to characterize the properties of uncovered representations
through computational modeling.
In Aim \#1, we will have collected tinnitus percept data using the reverse correlation method.
In Aim \#2, we will have reconstructed high-dimensional representations of tinnitus percepts
using both conventional and compressive sensing methods.
In this Aim, we will characterize the space of representations
to investigate correspondence between data-driven subtypes and qualitative subtypes,
evaluate the stability of tinnitus percepts over time,
and accelerate the diagnostic process.

\paragraph{Data-driven tinnitus subtypes}

Surveys of tinnitus patients have identified sixteen qualitative subtypes of tinnitus
\cite{vajsakovicPrinciplesMethodsPsychoacoustic2021,meikleTinnitusArchiveArchive2004,stoufferCharacterizationTinnitusTinnitus1990},
but these categories are limited by the fuzziness of linguistic categories and the musical ability of the tinnitus patients.
One tinnitus patient we interviewed, who has a degree in music performance,
was able to identify her tinnitus as sounding like a particular note played on the crotales,
a niche percussion instrument.
For most tinnitus patients who lack the depth of vocabulary to explain their tinnitus percept that precisely,
linguistic description of tinnitus can be vague and overlapping.
For instance, two of the aforementioned categories are ``high tension wire'' and ``transformer noise,''
which may mean different things to different people and are difficult to disambiguate.
Therefore, we propose to perform a quantitative analysis of the recovered high-dimensional tinnitus percepts
to reveal data-driven latent subtypes of tinnitus.

We will perform a quantitative analysis based on latent variable analysis methods.
Latent variable methods relate observed variation (\eg in tinnitus representations)
to latent variables that capture the observed variation (\eg major variation types across tinnitus percepts).
We will use principle component analysis (PCA), t-stochastic neighbor embedding (t-SNE; \cite{vandermaatenBarnesHutSNE2013}),
and uniform manifold approximation and projection (UMAP; \cite{mcinnesUMAPUniformManifold2020}).
Applying these dimensionality-reduction algorithms to a collection of representations
will identify major modes of variation across those representations,
and, in the case of PCA, provide an analysis of the proportion of variance explained by each mode
(the PCA eigenvalues).
Clusters of data points become readily visible in the lower-dimensional manifold generated
by these unsupervised machine learning methods, each of which represents
a data-driven tinnitus subtype.
We will characterize these subtypes, identifying their properties and assigning them descriptive names.
In the case of PCA and UMAP, new data can be mapped into the lower-dimensional manifold,
allowing representations from new subjects to be classified into these subtypes.

\paragraph{Stability of tinnitus percepts over time}

It is an open question in tinnitus research, whether tinnitus percepts remain stable over time
\cite{husainReplicabilityNeuralBehavioral2019}.
Using data from subjects who returned to perform the experiment multiple times
over the course of several weeks,
we will examine representations derived from each experiment instance
to quantitatively evaluate drift in the tinnitus representation over time.
We will assess measures of variance and spectral distance between representations
computed during different temporally-separated experiment sessions.

\paragraph{Acceleration of diagnosis through subtype identification}

While we anticipate remarkable performance gains in the number of samples required for accurate representation
through compressive sensing,
it may be clinically useful to identify only the subtype of tinnitus
(\ie which cluster on the lower-dimensional manifold the representation belongs to)
to some convergent accuracy.
If only the subtype is needed,
we believe the number of samples required for convergent accuracy can be further reduced.
We will devise an algorithm to perform the data-driven tinnitus subtyping ``online''
while the subject performs the task.
When convergent accuracy is reached, the task will automatically conclude.
By this process, both patient and clinician time is optimized, while still providing high-fidelity quantitative medical data
for use in diagnosis and treatment.
